Characterizing edges in signed and vector-valued graphs

### 2017-12-13 cc/bias_balance.tex
fix cc chapter layout

It is interesting to note that this local property of triangles gives raise to a global structure.
Namely, a strongly balanced complete graph can be split in two (possibly empty) clusters $L$ and $R$
such that all edges within $L$ and $R$ are positive, while all edges between $L$ and $R$ are
negative. Indeed, let us pick an arbitrary node $u$ and let $L=\neip(u)\setminus \{u\}$ and $R=V\setminus L$. Now
let $v,w\in L\setminus \{u\}$ and $x,y\in R$. From \autoref{fig:cc_twostrongcc} and the fact that
every triangle is balanced, we can see that $(v,w)$ and $(x,y)$ are positive edges, while $(w,y)$ is
a negative edge.
\begin{marginfigure}
  \centering
  \includegraphics[width=28mm]{assets/raw/2cc.pdf}
  \caption{A two-clustering of a complete strongly balanced graph}
  \label{fig:cc_twostrongcc}
\end{marginfigure}
We say that $L$ and $R$ are $2$-consistent clusters and more generally, we define
\begin{definition}[$K$-consistent clustering]
  Given a graph $G=(V, E)$, a clustering $\cluster=\{C_1, \ldots, C_K\}$ of $V$ is \emph{consistent}
  with the signs of $E$ if for every edge $(u,v)$ in $E$:
  \begin{enumerate}[(i),nosep]
    \item $(u,v)$ is positive if $\cluster(u)=\cluster(v)$, and
    \item $(u,v)$ is negative if $\cluster(u)\neq\cluster(v)$.
  \end{enumerate}
\end{definition}

>>>>

\begin{wrapfigure}{o}[10mm]{45mm}
  \centering
  \includegraphics[width=.9\linewidth]{assets/raw/2cc.pdf}
  \caption{\small A two-clustering of a complete strongly balanced graph}
  \label{fig:cc_twostrongcc}
\end{wrapfigure}
It is interesting to note that this local property of triangles gives raise to a global structure.
Namely, a strongly balanced complete graph can be split in two (possibly empty) clusters $L$ and $R$
such that all edges within $L$ and $R$ are positive, while all edges between $L$ and $R$ are
negative. Indeed, let us pick an arbitrary node $u$ and let $L=\neip(u)\setminus \{u\}$ and $R=V\setminus L$. Now
let $v,w\in L\setminus \{u\}$ and $x,y\in R$. From \autoref{fig:cc_twostrongcc} and the fact that
every triangle is balanced, we can see that $(v,w)$ and $(x,y)$ are positive edges, while $(w,y)$ is
a negative edge.
We say that $L$ and $R$ are $2$-consistent clusters and more generally, we define
\begin{definition}[$K$-consistent clustering]
  Given a graph $G=(V, E)$, a clustering $\cluster=\{C_1, \ldots, C_K\}$ of $V$ is \emph{consistent}
  with the signs of $E$ if for every edge $(u,v)$ in $E$:
  \begin{enumerate}[(i),nosep]
    \item $(u,v)$ is positive if $\cluster(u)=\cluster(v)$, and
    \item $(u,v)$ is negative if $\cluster(u)\neq\cluster(v)$.
  \end{enumerate}
\end{definition}

### 2017-12-13 cc/gtx_algorithm.tex
fix cc chapter layout

\paragraph{\extractStar{}}\label{par:extractstar}%
\extractStar{} takes as input a graph $G_t=(V_t, E_t)$.
While the nodeset $V_t$ is not exhausted, it repeatedly samples a
node $c_i$, creates a star $S_i^t$ with $c_i$ at its center and the neighbors of $c_i$ on the
periphery, removes all the nodes of $S_i^t$ from
$V_t$ and all the edges incident to $S_i^t$ from $E_t$, and finally decrements accordingly the
degree of the 2-hop neighbors of $c_i$ (see \autoref{fig:gtx_star_simple} for a visual
representation of this notation).
Upon completion, it returns a list of stars,
the set of all the edges within a star,
and a map (or associative array) that associates each node of $V_t$ to the index of the
unique star it belongs to.
According to the definition of \textcite{HashTableBook08}, an associative array is an abstract data
type composed of a collection of (key, value) pairs, such that each possible key appears at most
once in the collection. It efficiently supports the addition, removal and modification of a pair, as
well as the lookup of a value associated with a particular key.
\begin{marginfigure}
  \centering
  \includegraphics[height=0.15\textheight]{assets/tikz/gtx_star_tikz.pdf}
  \caption[A sample star]{A sample star created during the \tth{} collapse level. The black node
    % \tikz{\node[vertex,rare] {$c_i$};}
    is the center $c_i$ of the star $S_i^t$, which is also made of the four light gray peripheral nodes
  % \tikz{\node[vertex,medium] {$p_1$};} to \tikz{\node[vertex,medium] {$p_4$};}
  as well as the solid edges. The 2-hops neighbors of $c_i$ are the white nodes
  % \tikz{\node[vertex] {$h_1$};} to \tikz{\node[vertex] {$h_3$};}
  $h_1$ to $h_3$, whose degree will decrease once $S_i^t$ is removed from $G_t$.}
  \label{fig:gtx_star_simple}
\end{marginfigure}

@@

\begin{center}
  \rule{\textwidth}{.3pt}
  \begin{algorithmic}[1]
    \Function{\extractStar{}}{$G_t=(V_t,E_t)$}
      \State Let $Q$ be the max-priority queue described above
      \State Let $remaining$ be a set of nodes, initially containing all the nodes in $V_t$
      \State Let $membership$ be an empty map
      \Let{$stars$}{$[]$}
      \Let{$inner\_edges$}{$\emptyset$}
      \While{$Q$ is not empty}
        \Let{$c$}{\Call{Extract-Max}{$Q$}}
        \If{$c$ not in $remaining$}
          \State \textbf{continue} \Comment{$c$ is part of an existing star so there is
          nothing to do}
        \EndIf
        \Let{$periphery$}{$G_t[c] \bigcap remaining$}
        \Let{$stars$}{$stars \bigcup \{$\Call{Star}{$c$, $periphery$, $|stars|$}\}}
        \Let{$inner\_edges$}{$inner\_edges \bigcup \{(c, p): p \in periphery\}$}
        \Let{$remaining$}{$remaining \setminus \left\{ \{c_i\} \cup periphery\right\}$}
        \For{$p$ in $periphery$}
          \For{$h$ in $G_t[p] \bigcap remaining$}
            \State \Call{Decrease-Key}{$Q$, $h$, $1$}
          \EndFor
        \EndFor
      \EndWhile
      \State \textbf{return} $stars$, $inner\_edges$, $membership$
    \EndFunction
  \end{algorithmic}
  \rule{\textwidth}{.3pt}
\end{center}

\begin{prop}
  \label{prop:gtx_time_extractstar}
  For any connected graph $G=(V,E)$, \extractStar$(G)$ terminates in $O(|E|)$ time.
\end{prop}
\begin{proof}
\extractStar{} terminates because at each iteration of the while loop line 7, we remove one node
from $Q$ and never add any. Let us now analyze its complexity. We first build a priority queue of
all the nodes according to their degree (line 2), which requires $|V|$ insertions into $Q$. Then
we execute $|V|$ iterations of the while loop. However, the main idea here is that we process each
node and each edge exactly once. We first find the center $c$ of the next star by extracting the
maximum of the queue (line 8) and testing if $c$ is still part of the graph, which happens $|V|$
times in total. Then we build the corresponding star (line 11--14). In total, we test the membership
of $|V|$
nodes in line 11, $\textsc{Star}$ updates the $membership$ map $|V|$ times in line 12,
$inner\_edges$ consists of $|E|$ edges at most in line 13 and $remaining$ is only updated $|V|$
times in line 14. Finally, we decrease the priority (\ie{} the degree) of all nodes adjacent to the
new star (line 15--17). Each decrement is supported by a single edge, thus \textsc{Decrease-Keys} is
called at most $|E|$ times. Since all queue operations require constant time when using a Strict
Fibonacci Heap~\autocite{FibonacciHeaps12}, the complexity is $O(|E|+|V|)$, which is also $O(|E|)$
since $G$ is connected.
\end{proof}

@@

\begin{marginfigure}
  \centering
  \includegraphics[width=0.92\linewidth]{assets/tikz/gtx_cross_edges_tikz.pdf}
  \caption[Cross edge representation]{In this graph, there are four possible edges between the two
  stars $s_1$ and $s_2$, and all are part of $cross\_edges[(s_1, s_2)]$, here represented in light
red. Those edges are labeled with the sum of eccentricity of their underlying endpoints. White node
have an eccentricity of $0$, and gray node an eccentricity of $1$. The edge with the
minimal eccentricity is linking a peripheral node of $s_1$ directly to the center of $s_2$.}
  \label{fig:gtx_cross_edge}
\end{marginfigure}%
\collapseStar{} not only returns the new graph $G_{t+1}$ but also a map $cross\_edges$. This map is
used to keep track of which edge in $E_t$ connects any pair of node in $V_{t+1}$ (as illustrated in
\autoref{fig:gtx_cross_edge}). It associates to any edge in $E_{t+1}$ a set of edges from $E_t$.
This set can have an arbitrary size during the execution of \collapseStar{}, yet it is guaranteed to
contain a single edge upon termination. Although this does not appear in the following pseudo code,
in practice we ensure that $(s_u, s_v)$ and $(s_v, s_u)$ refer to the same edge in $cross\_edges$.

\begin{center}
  \rule{\textwidth}{.3pt}
  \begin{algorithmic}[1]
    \Function{\collapseStar{}}{$E_t,\,membership,\,eccentricity$}
      \State Let $G_{t+1}$ be an empty graph
      \State Let $cross\_edges$ be the map described above
      \For{\textbf{every} edge $(u, v)$ in $E_t$}
      \Let{$s_u$}{$membership[u]$}
      \Let{$s_v$}{$membership[v]$}
        \If{$u$ and $v$ are not in the same star (\ie{} $s_u \neq s_v$)}
          \Let{$cross\_edges[(s_u, s_v)]$}{$cross\_edges[(s_u, s_v)] \bigcup \{(u, v)\}$}
        \EndIf
      \EndFor
      \For{\textbf{every} pair of will-be connected stars $(s_u,s_v)$ in $cross\_edges$}
        \Let{$possible\_underlying\_edges$}{$cross\_edges[(s_u, s_v)]$}
        %\marginpars{\small Line 11 is actually a simplification since the $eccentricity$ array is
        %only defined for nodes in $V_0$ so when \collapseStar{} is called on \eg{} $G_3$, $u$
        %and $v$ are nodes of $V_3$ and we have to retrieve the actual corresponding edge in $E_0$.
        %Maybe I could simply add a map from $E_t$ to $E_0$ like in the real python code…}
        \State $$(u_0, v_0) \gets \argmin_{(u,v)\in possible\_underlying\_edges} eccentricity[u]+eccentricity[v]$$
        \Let{$cross\_edges[(s_u, s_v)]$}{$\{(u_0, v_0)\}$}
        \Let{$E_{t+1}$}{$E_{t+1} \bigcup \{(s_u, s_v)\}$}
      \EndFor
      \State \textbf{return} $G_{t+1}$, $cross\_edges$
    \EndFunction
  \end{algorithmic}
  \rule{\textwidth}{.3pt}
\end{center}

@@

\vspace{0.5\baselineskip}
\setlength{\fullpage}{\textwidth+\marginparsep+\marginparwidth}
\noindent\begin{tabulary}{\fullpage}{lllL}
  \toprule
  name               & keys set at $t$ & values set at $t$   & comment \\
  \midrule
  $star\_membership$ & nodes in $V_t$  & nodes in $V_{t+1}$  & this is the map returned by
  \extractStar{}$(G_t)$ \\
  $full\_membership$ & nodes in $V_0$  & nodes in $V_{t+1}$  & at $t=0$, this is equals to
  $star\_membership$ and then it gets updated at every iteration to maintain its original keys set.  \\
  $original\_node$   & nodes in $V_t$  & sets of $V_0$ nodes & This can be seen as the reverse of
  $full\_membership$ and it is the one needed to update the eccentricity of the original nodes. \\
  \bottomrule
\end{tabulary}
\vspace{0.5\baselineskip}

@@

  Finally we prove that \gtx{} indeed returns a spanning tree of $G_0$. Specifically, we show that
  by starting from $G_K$ and going in reverse order, we build a spanning tree $T_t$ of every $G_t$,
  including eventually $G_0$. This process is illustrated on \autoref{fig:gtx_spanning_tree} and
  works as follow: when moving from $G_{t+1}$ to $G_t$, we expand the nodes of $G_{t+1}$ into star
  of nodes of $V_t$ and update all the outer edges we met so far to have their endpoints in $V_t$.
  These outer edges are exactly $T_{t+1}$ and by translating them into $E_t$ and adding the inner
  star edges of $G_t$, we build $T_t$. $G_K$ consists of a single node with no edges so we have
  trivially that $T_K$ = $G_K$. Then, assume we have a spanning tree $T_{t+1}$ of $G_{t+1}$, so that
  all nodes in $V_{t+1}$ are connected without cycle. Expanding each node of $V_{t+1}$ into a star
  of $V_t$ nodes ensures that the spanning property is preserved since each node of $V_t$ is covered
  by construction. Those stars are also without cycle by construction. The second step, translating
  the existing edges of $T_{t+1}$, only connects nodes belonging to different stars and because
  those edges were cycle free in $G_{t+1}$, this remains the case in $G_t$. Therefore we have build
  a spanning tree $T_t$ of $G_t$. By repeating this procedure, the \gtx{} algorithm eventually
  builds a spanning tree of $G_0$.
\end{proof}
\vspace{-2\baselineskip}
\begin{figure*}[h]
  \centering
	\begin{subfigure}[b]{37mm}
    \centering
		\includegraphics[width=\textwidth]{assets/tikz/gtx_st_g3_tikz.pdf}
		\caption{Spanning tree of $G_3$}\label{fig:gtx_spanning_tree_3}
	\end{subfigure}~
	\begin{subfigure}[b]{45mm}
    \centering
		\includegraphics[width=\textwidth]{assets/tikz/gtx_st_g2_tikz.pdf}
		\caption{Spanning tree of $G_2$}\label{fig:gtx_spanning_tree_2}
	\end{subfigure}~
	\begin{subfigure}[b]{49mm}
    \centering
		\includegraphics[width=\textwidth]{assets/tikz/gtx_st_g1_tikz.pdf}
		\caption{Spanning tree of $G_1$}\label{fig:gtx_spanning_tree_1}
	\end{subfigure}~
	\begin{subfigure}[b]{53mm}
    \centering
		\includegraphics[width=\textwidth]{assets/tikz/gtx_st_g0_tikz.pdf}
		\caption{Spanning tree of $G_0$}\label{fig:gtx_spanning_tree_0}
	\end{subfigure}~
  \caption{Unfolding stars to recover spanning trees}
  \label{fig:gtx_spanning_tree}
\end{figure*}

@@

A crucial quantity of the \gtx{} algorithm, both in terms of complexity and resulting stretch, is
the number of collapses $K$ needed before termination. While this is still elusive to express in the
general case, let us first look at some simple cases. For instance, a very sparse example of graph
is the line graph. While it is already a tree, let us look how the \gtx{} algorithm operates on it.
Say we have $n$ nodes and $m$ edges in that line. Here, a star is made at most of three
consecutive nodes. In the worst case, centers will be chosen such that we have a succession of three
and two nodes stars, as in \autoref{fig:gtx_line_graph}.%
\begin{marginfigure}
  \centering
  \includegraphics[width=0.9\linewidth]{assets/tikz/gtx_line_tikz.pdf}
  \caption{A line graph with stars in blue.}
  \label{fig:gtx_line_graph}
\end{marginfigure}
This will results in $\nicefrac{2n}{5}$ stars organized in a line, which is less than half of $n$.
Thus we can conclude that in this case, \gtx{} will finish after $O(\log n)$ iterations. Note that a barbell graph (two
cliques connected by a line) would requires a number of iterations logarithmic in the length of that
central line, despite having many more nodes and edges.

>>>>

\paragraph{\extractStar{}}\label{par:extractstar}%
\begin{wrapfigure}{o}[10mm]{45mm}
  \centering
  \includegraphics[width=.9\linewidth]{assets/tikz/gtx_star_tikz.pdf}
  \caption[A sample star]{\small A sample star created during the \tth{} collapse level. The black node
    is the center $c_i$ of the star $S_i^t$, which is also made of the four light gray peripheral nodes
  as well as the solid edges. The 2-hops neighbors of $c_i$ are the white nodes
  $h_1$ to $h_3$, whose degree will decrease once $S_i^t$ is removed from $G_t$.}
  \label{fig:gtx_star_simple}
\end{wrapfigure}
\extractStar{} takes as input a graph $G_t=(V_t, E_t)$.
While the nodeset $V_t$ is not exhausted, it repeatedly samples a
node $c_i$, creates a star $S_i^t$ with $c_i$ at its center and the neighbors of $c_i$ on the
periphery, removes all the nodes of $S_i^t$ from
$V_t$ and all the edges incident to $S_i^t$ from $E_t$, and finally decrements accordingly the
degree of the 2-hop neighbors of $c_i$ (see \autoref{fig:gtx_star_simple} for a visual
representation of this notation).
Upon completion, it returns a list of stars,
the set of all the edges within a star,
and a map (or associative array) that associates each node of $V_t$ to the index of the
unique star it belongs to.
According to the definition of \textcite{HashTableBook08}, an associative array is an abstract data
type composed of a collection of (key, value) pairs, such that each possible key appears at most
once in the collection. It efficiently supports the addition, removal and modification of a pair, as
well as the lookup of a value associated with a particular key.

@@

\begin{center}
  \rule{\textwidth}{.3pt}
  \begin{algorithmic}[1]
    \Function{\extractStar{}}{$G_t=(V_t,E_t)$}
      \State Let $Q$ be the max-priority queue described above
      \State Let $remaining$ be a set of nodes, initially containing all the nodes in $V_t$
      \State Let $membership$ be an empty map
      \Let{$stars$}{$[]$}
      \Let{$inner\_edges$}{$\emptyset$}
      \While{$Q$ is not empty}
        \Let{$c$}{\Call{Extract-Max}{$Q$}}
        \If{$c$ not in $remaining$}
          \State \textbf{continue} \Comment{$c$ is part of an existing star so there is
          nothing to do}
        \EndIf
        \Let{$periphery$}{$G_t[c] \bigcap remaining$}
        \Let{$stars$}{$stars \bigcup \{$\Call{Star}{$c$, $periphery$, $|stars|$}\}}
        \Let{$inner\_edges$}{$inner\_edges \bigcup \{(c, p): p \in periphery\}$}
        \Let{$remaining$}{$remaining \setminus \left\{ \{c_i\} \cup periphery\right\}$}
        \For{$p$ in $periphery$}
          \For{$h$ in $G_t[p] \bigcap remaining$}
            \State \Call{Decrease-Key}{$Q$, $h$, $1$}
          \EndFor
        \EndFor
      \EndWhile
      \State \textbf{return} $stars$, $inner\_edges$, $membership$
    \EndFunction
  \end{algorithmic}
  % \rule{\textwidth}{.3pt}
\end{center}

\vspace*{-2\baselineskip}
\begin{prop}
  \label{prop:gtx_time_extractstar}
  For any connected graph $G=(V,E)$, \extractStar$(G)$ terminates in $O(|E|)$ time.
\end{prop}
\begin{proof}
\extractStar{} terminates because at each iteration of the while loop line 7, we remove one node
from $Q$ and never add any. Let us now analyze its complexity. We first build a priority queue of
all the nodes according to their degree (line 2), which requires $|V|$ insertions into $Q$. Then
we execute $|V|$ iterations of the while loop. However, the main idea here is that we process each
node and each edge exactly once. We first find the center $c$ of the next star by extracting the
maximum of the queue (line 8) and testing if $c$ is still part of the graph, which happens $|V|$
times in total. Then we build the corresponding star (line 11--14). In total, we test the membership
of $|V|$
nodes in line 11, $\textsc{Star}$ updates the $membership$ map $|V|$ times in line 12,
$inner\_edges$ consists of $|E|$ edges at most in line 13 and $remaining$ is only updated $|V|$
times in line 14. Finally, we decrease the priority (\ie{} the degree) of all nodes adjacent to the
new star (line 15--17). Each decrement is supported by a single edge, thus \textsc{Decrease-Keys} is
called at most $|E|$ times. Since all queue operations require constant time when using a Strict
Fibonacci Heap~\autocite{FibonacciHeaps12}, the complexity is $O(|E|+|V|)$, which is also $O(|E|)$
since $G$ is connected.
\end{proof}

@@

\collapseStar{} not only returns the new graph $G_{t+1}$ but also a map $cross\_edges$. This map is
used to keep track of which edge in $E_t$ connects any pair of node in $V_{t+1}$ (as illustrated in
\autoref{fig:gtx_cross_edge}). It associates to any edge in $E_{t+1}$ a set of edges from $E_t$.
This set can have an arbitrary size during the execution of \collapseStar{}, yet it is guaranteed to
contain a single edge upon termination. Although this does not appear in the following pseudo code,
in practice we ensure that $(s_u, s_v)$ and $(s_v, s_u)$ refer to the same edge in $cross\_edges$.

\begin{figure}[bpt]
  \centering
  \includegraphics[width=0.4\linewidth]{assets/tikz/gtx_cross_edges_tikz.pdf}
  \caption[Cross edge representation]{In this graph, there are four possible edges between the two
  stars $s_1$ and $s_2$, and all are part of $cross\_edges[(s_1, s_2)]$, here represented in light
red. Those edges are labeled with the sum of eccentricity of their underlying endpoints. White node
have an eccentricity of $0$, and gray node an eccentricity of $1$. The edge with the
minimal eccentricity is linking a peripheral node of $s_1$ directly to the center of $s_2$.}
  \label{fig:gtx_cross_edge}
\end{figure}

\begin{center}
  \rule{\textwidth}{.3pt}
  \begin{algorithmic}[1]
    \Function{\collapseStar{}}{$E_t,\,membership,\,eccentricity$}
      \State Let $G_{t+1}$ be an empty graph
      \State Let $cross\_edges$ be the map described above
      \For{\textbf{every} edge $(u, v)$ in $E_t$}
      \Let{$s_u$}{$membership[u]$}
      \Let{$s_v$}{$membership[v]$}
        \If{$u$ and $v$ are not in the same star (\ie{} $s_u \neq s_v$)}
          \Let{$cross\_edges[(s_u, s_v)]$}{$cross\_edges[(s_u, s_v)] \bigcup \{(u, v)\}$}
        \EndIf
      \EndFor
      \For{\textbf{every} pair of will-be connected stars $(s_u,s_v)$ in $cross\_edges$}
        \Let{$possible\_underlying\_edges$}{$cross\_edges[(s_u, s_v)]$}
        %\marginpars{\small Line 11 is actually a simplification since the $eccentricity$ array is
        %only defined for nodes in $V_0$ so when \collapseStar{} is called on \eg{} $G_3$, $u$
        %and $v$ are nodes of $V_3$ and we have to retrieve the actual corresponding edge in $E_0$.
        %Maybe I could simply add a map from $E_t$ to $E_0$ like in the real python code…}
        \State $$(u_0, v_0) \gets \argmin_{(u,v)\in possible\_underlying\_edges} eccentricity[u]+eccentricity[v]$$
        \Let{$cross\_edges[(s_u, s_v)]$}{$\{(u_0, v_0)\}$}
        \Let{$E_{t+1}$}{$E_{t+1} \bigcup \{(s_u, s_v)\}$}
      \EndFor
      \State \textbf{return} $G_{t+1}$, $cross\_edges$
    \EndFunction
  \end{algorithmic}
  \vspace*{-.5\baselineskip}
  % \rule{\textwidth}{.3pt}
\end{center}

@@

\vspace{0.5\baselineskip}
\setlength{\fullpage}{179mm}
\begin{center}
\begin{adjustwidth}{-2cm}{}
\noindent\begin{tabulary}{\fullpage}{lllL}
  \toprule
  name               & keys set at $t$ & values set at $t$   & comment \\
  \midrule
  $star\_membership$ & nodes in $V_t$  & nodes in $V_{t+1}$  & this is the map returned by
  \extractStar{}$(G_t)$ \\
  $full\_membership$ & nodes in $V_0$  & nodes in $V_{t+1}$  & at $t=0$, this is equals to
  $star\_membership$ and then it gets updated at every iteration to maintain its original keys set.  \\
  $original\_node$   & nodes in $V_t$  & sets of $V_0$ nodes & This can be seen as the reverse of
  $full\_membership$ and it is the one needed to update the eccentricity of the original nodes. \\
  \bottomrule
\end{tabulary}
\end{adjustwidth}
\end{center}
\vspace{0.5\baselineskip}

@@

  Finally we prove that \gtx{} indeed returns a spanning tree of $G_0$. Specifically, we show that
  by starting from $G_K$ and going in reverse order, we build a spanning tree $T_t$ of every $G_t$,
  including eventually $G_0$. This process is illustrated on \autoref{fig:gtx_spanning_tree} and
  works as follow: when moving from $G_{t+1}$ to $G_t$, we expand the nodes of $G_{t+1}$ into star
  of nodes of $V_t$ and update all the outer edges we met so far to have their endpoints in $V_t$.
  These outer edges are exactly $T_{t+1}$ and by translating them into $E_t$ and adding the inner
  star edges of $G_t$, we build $T_t$. $G_K$ consists of a single node with no edges so we have
  trivially that $T_K$ = $G_K$. Then, assume we have a spanning tree $T_{t+1}$ of $G_{t+1}$, so that
  all nodes in $V_{t+1}$ are connected without cycle. Expanding each node of $V_{t+1}$ into a star
  of $V_t$ nodes ensures that the spanning property is preserved since each node of $V_t$ is covered
  by construction. Those stars are also without cycle by construction. The second step, translating
  the existing edges of $T_{t+1}$, only connects nodes belonging to different stars and because
  those edges were cycle free in $G_{t+1}$, this remains the case in $G_t$. Therefore we have build
  a spanning tree $T_t$ of $G_t$. By repeating this procedure, the \gtx{} algorithm eventually
  builds a spanning tree of $G_0$.
\end{proof}
\vspace{-2\baselineskip}
\begin{figure*}[h]
  \centering
	\begin{subfigure}[b]{37mm}
    \centering
		\includegraphics[width=\textwidth]{assets/tikz/gtx_st_g3_tikz.pdf}
		\caption{Spanning tree of $G_3$}\label{fig:gtx_spanning_tree_3}
	\end{subfigure}~
	\begin{subfigure}[b]{45mm}
    \centering
		\includegraphics[width=\textwidth]{assets/tikz/gtx_st_g2_tikz.pdf}
		\caption{Spanning tree of $G_2$}\label{fig:gtx_spanning_tree_2}
	\end{subfigure}~
	\begin{subfigure}[b]{49mm}
    \centering
		\includegraphics[width=\textwidth]{assets/tikz/gtx_st_g1_tikz.pdf}
		\caption{Spanning tree of $G_1$}\label{fig:gtx_spanning_tree_1}
	\end{subfigure}

	\begin{subfigure}[b]{53mm}
    \centering
		\includegraphics[width=\textwidth]{assets/tikz/gtx_st_g0_tikz.pdf}
		\caption{Spanning tree of $G_0$}\label{fig:gtx_spanning_tree_0}
	\end{subfigure}~
  \caption{Unfolding stars to recover spanning trees}
  \label{fig:gtx_spanning_tree}
\end{figure*}

@@

A crucial quantity of the \gtx{} algorithm, both in terms of complexity and resulting stretch, is
the number of collapses $K$ needed before termination. While this is still elusive to express in the
general case, let us first look at some simple cases. For instance, a very sparse example of graph
is the line graph. While it is already a tree, let us look how the \gtx{} algorithm operates on it.
Say we have $n$ nodes and $m$ edges in that line. Here, a star is made at most of three
consecutive nodes. In the worst case, centers will be chosen such that we have a succession of three
and two nodes stars, as in \autoref{fig:gtx_line_graph}.%
This will results in $\nicefrac{2n}{5}$ stars organized in a line, which is less than half of $n$.
Thus we can conclude that in this case, \gtx{} will finish after $O(\log n)$ iterations. Note that a barbell graph (two
cliques connected by a line) would requires a number of iterations logarithmic in the length of that
central line, despite having many more nodes and edges.
\begin{figure}[bth]
  \centering
  \includegraphics[width=0.9\linewidth]{assets/tikz/gtx_line_tikz.pdf}
  \caption{A line graph with stars in blue.}
  \label{fig:gtx_line_graph}
\end{figure}
